---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup_knitr_options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# fishcastr

<!-- badges: start -->
[![R-CMD-check](https://github.com/as-french/fishcastr/workflows/R-CMD-check/badge.svg)](https://github.com/as-french/fishcastr/actions)
<!-- badges: end -->

# What is fishcastr?

`fishcastr` is an R package that provides a collection of tools for generating seasonal migration forecasts for the timing of seaward movements of diadromous fishes.

# Why forecast diadromous fish movements?

Diadromous fish migration forecasts can inform management of in-river infrastructures that impede fish movement. Globally, the seaward movements of diadromous fishes are impeded by anthropogenic in-river infrastructure (e.g., hydropower generation turbines, flood management pump stations and dams). To facilitate seaward movement, infrastructure operators employ measures to reduce direct mortality, such as trap and transport or temporarily shutting down power generation. The success of these measures depend on being able to predict the times during which the largest cohorts of migrating fish will be captured. Advanced warning of migration timing (e.g., onset and cessation of migration periods) therefore has the potential to improve the survival of vulnerable populations.

# What is a seasonal migration forecast?

Seasonal migration forecasts are communicated in terms of probabilistic predictions of timing anomalies (e.g., early, late or average mean date of migration). For consistency with other seasonal environmental forecasts, seasonal migration forecast are produced by applying multi-member ensemble *seasonal climate forecast* forcing data, such as [ECMWF's SEAS5](https://gmd.copernicus.org/articles/12/1087/2019/) product, to a model chain. This application produces an ensemble of plausible daily count simulations/scenarios within a target season of interest (e.g., March through July). From these ensemble simulations, migration timing statistics (e.g., mean date of migration) are calculated for each member and ranked by tercile frequencies; i.e., 1/25 members predict lower tercile (early migration), 20/25 predict upper tercile (late migration), and 4/25 predict middle tercile (average migration).

# Why develop seasonal forecasts for diadromous fishes?
 
The potential of *seasonal climate forecasts* for informing environmental and ecological management through operational forecasts is receiving increasing attention (e.g., in marine and freshwater settings, as well as fire risk).

Diadromous fish migrations occur seasonally (e.g., the seaward spring Atlantic salmon (*Salmo salar*) juvenile (smolt) migration), and variability in timing among catchments associates with latitude, but also varies inter-annually within catchments. This inter-annual variability within catchments is associated with seasonal climate variability at short and medium time scales. In Europe, for example, movements of migrating Atlantic salmon (*Salmo salar*), andadromous brown trout (*Salmo trutta*) and European eel (*Anguilla anguilla*) are recorded by a sparse network of in-river trapping facilities where fish counts associate with daily environmental variability related to recent weather (e.g., river discharge and temperature), and other factors associated with predator avoidance (e.g., moonlight exposure). Moreover, prior to sea entry, diadromous fishes undergo a physiological transition that prepare them for marine endurance (e.g., the smoltification transition from resident juvenile salmon parr to migratory salmon smolt), and the initialisation and time it takes for this transition to complete is associated with seasonal climate variability that can be quantified (for example) using degree days. If modelled, interactions among environmental and climate variables associated with environmental tolerability for movement and physiological preparedness for sea entry can be used to predict inter-annual variation in the timing of migration of diadromous fishes. The added value to seasonal predictions is the advanced warning they can provide for those who are responsible for the logistics such as trap and transport and power generation contingency plans.

# What does fishcastr do?

`fishcastr`'s functionality is demonstrated in the package vignettes. The vignettes "manuscript" and "supplementary materials" describe the process of accessing and pre-process publicly available fish count data, environmental data and climate data and use open-source statistical and process based modelling tools to build a forecast workflow and evaluate the credibility of its predictions by retrospective skill assessment.

## A non-exhaustive list of functionality

- Generate ensemble fish count simulations from seasonal climate forecast input data and a model chain (see below)
- Convert ensemble simulations to probabilistic migration timing forecasts for a range of seasonal timing summary statistics (mean day of migration, percentiles etc.)
- Visualise uncertainty in probabilistic predictions, building on some of the functionality in the [visualizeR](https://github.com/SantanderMetGroup/visualizeR) package for seasonal climate forecast visualization
- Predict water temperature from air temperature
- Calculate moonlight exposure
- Stratify count census data into "forecast years" based on biological information (thermal/day length thresholds)

***Model chain***

The example seasonal migration forecasts contained in the `fishcastr` vignettes are derived from a model chain that includes a process based rainfall run-off model (using [airGR](https://github.com/cran/airGR)) and correlative fish count models (using [glmmTMB](https://github.com/glmmTMB/glmmTMB)). Additional elements of the model chain, including a statistical water temperature model and moonlight exposure estimates are implemented within `fishcastr`.

# How to install

Install Rtools:
https://cran.r-project.org/bin/windows/Rtools/

Install R package:

```{r how_to_install, include = TRUE,eval=FALSE, echo=TRUE}
install.packages("devtools")
devtools::install_github("as-french/fishcastr",
                         build_vignettes = FALSE)
```


## A note on access to archived seasonal climate forecasts and grid pseudo-observations

`fishcastr` includes all data required to reproduce the analysis described in the "vignettes/Manuscript". For example, re-forecasts and archived operational forecasts for 1993 to 2019 interpolated to the location of the Burrishoole catchment in the west of Ireland can be found in raw form (`fishcastr::grid_SEAS5_1993_2019_2_3_4_5_6_7_8_tas_pr_petH_raw`) and bias corrected form (`fishcastr::grid_SEAS5_1993_2019_2_3_4_5_6_7_8_tas_pr_petH_bc`) - though note the values in both these datasets are rounded to reduce package size.

For users who wish to access raw seasonal climate forecast data for other locations, there are two options.

- Contact the University of Santander Met Group for access to archived SEAS5 forecasts through the [Climate4R](https://github.com/SantanderMetGroup/climate4R) package bundle using functions such as `loadeR::loadSeasonalForecast`, or..

- Download climate data directly from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/#!/home) in netcdf format. If you choose this option, we recommend using the [ecmwfr](https://cran.r-project.org/web/packages/ecmwfr/vignettes/webapi_vignette.html) package for data download/documentation, and we recommend you use the `Climate4R` tools for data processing. The /raw-data/ folder in this package contains example scripts for accessing ECMWF's ERA5 reanalysis and operational SEAS5 data from Copernicus CDS.

In both instances, we recommend installation of the `Climate4R` bundle - see instructions on each github page, taking particular care with loadeR, owing to its dependency on a Java installation and the R package rJava, which has [known issues](https://github.com/SantanderMetGroup/loadeR/wiki/Installation).

Install in following order:

- [rJava](https://cran.r-project.org/web/packages/rJava/index.html)
- [loadeR.java](https://github.com/SantanderMetGroup/loadeR.java)
- [climate4R.UDG](https://github.com/SantanderMetGroup/climate4R.UDG)
- [loadeR](https://github.com/SantanderMetGroup/loadeR)

```{r how_to_install_Climate4R_packages, include = TRUE,eval=FALSE, echo=TRUE}
# ----------------------------------------------------------------------------------- #
# SETTING UP R FOR fishcastr PACKAGE INSTALLATION (FOR WINDOWS 10)
# ----------------------------------------------------------------------------------- #

# NOTES ----
# Package built using R version 3.6.3 on windows 10
# Might need to disable some aspect of anti-virus software to allow packages to be installed (e.g., "Safe Files" for Bitdefender 2018v onwards)

# DOWNLOAD AND INSTALL Java ----
# Download java 12.0.2 jdk 64 bit (note version 14 is not compatible; date 19-09-2020)
# https://www.oracle.com/java/technologies/javase/jdk12-archive-downloads.html
# see https://www.oracle.com/java/technologies/javase-downloads.html for other versions

# SET NEW ENVIRONMENT VARIABLE "JAVA_HOME" ----
# either manually by adding to environment path JAVA_HOME: C:\Program Files\Java\jdk-12.0.2 
# or set in r using code
# Sys.setenv(JAVA_HOME="C:/Program Files/Java/jdk-12.0.2/")

# ADD JDK bin PATH TO SYSTEM ENV PATH ----
# C:\Program Files\Java\jdk-12.0.2\bin to system Path variable

# INSTALL Rtools35 ----
# https://cran.r-project.org/bin/windows/Rtools/history.html

# INSTALL devtools R package ----
# install.packages("devtools")

# INSTALL rJava R package ----
# install.packages("rJava",type = "source")

# INSTALL UNICAN R packages ----
# NOTE - this is not necessary for fishcastr to install, or for vignette
# functions to work, but is necessary for developing forecasts for other regions
# following protocol contained in the Manuscript vignette
#### noting we have installed 64-bit java only; hence "--no-multiarch"
# devtools::install_github("SantanderMetGroup/loadeR.java", INSTALL_opts=c("--no-multiarch"))
# devtools::install_github("SantanderMetGroup/climate4R.UDG", INSTALL_opts=c("--no-multiarch"))
# devtools::install_github("SantanderMetGroup/loadeR", INSTALL_opts=c("--no-multiarch"))
# devtools::install_github(c("SantanderMetGroup/transformeR",
#                            "SantanderMetGroup/downscaleR",
#                            "SantanderMetGroup/visualizeR",
#                            "SantanderMetGroup/drought4R"))
```

# How to use

See vignettes.

# Issues

This package is maturing and has many areas that could be improved. If you find an issue, please raise it at: https://github.com/as-french/fishcastr/issues.

# Credits

`fishcastr` was built to support the ClimateJPI ERA4CS WATExR project 2017 - 2021 (https://watexr.eu).

# Literature cited
